---
title: "Statistical Computing Final Project"
author: "MustafaCanİnce-200709081"
date: "2023-06-14"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. (3p) Data:

Please find your original dataset or datasets; and describe your data in
the first step

As real data, the data of the series and movies I watched, my rating(0
to 10), the date I gave the score, title, Url, imdb average (0 to 10),
runtime(min), release year, genres, Total number of votes , Released
year, Director.

```         
```

## 2. (3p) Exploratory and descriptive data analysis:

Use "Exploratory and descriptive data analysis". Talk about your
categorical and quantitative data or your ordinal variables etc. Write
down your comments

```{r pressure, echo=FALSE}
required_packages <- c("rstudioapi", "ggplot2", "dplyr", "scales")

for (package in required_packages) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
  if (require(package, character.only = TRUE)) {
    library(package, character.only = TRUE)
  }
}

my_data <- read.csv('ratings.csv')


```

Const : ('tt1014759', tt0101540, ...)
Your.Rating: ('5', '8', ...)
Date.Rated: ('1/16/2020', '9/12/2022', ...), (mm, dd, yyyy)
Title: ('Alice in Wonderland', 'Cape Fear', ...)
URL: ('https://www.imdb.com/title/tt1014759/', 'https://www.imdb.com/title/tt0101540/', ...)
Title.Type: ('movie', 'movie', ...)
IMDb.Rating: ('6.4', '7.3', ...)
Runtime..mins.: ('108', '128', ...)
Year: ('2010', '1991', ...)
Genres: ('Adventure, Family, Fantasy, Mystery', 'Crime, Thriller', ...)
Num.Votes: ('426039', '204632', ...)
Release.Date: ('2/25/2010', '10/6/1991', ...), (mm, dd, yyyy)
Directors: ('Tim Burton', 'Martin Scorsese', ...)

Categorical Variables:
'Const', 'Title', 'Title.Type', 'Genres'

Ordinal Variables:
'Your.Rating'

Interval Variables:
'Date.Rated', 'Runtime..mins.', 'Year', 'Num.Votes', 'Release.Date'

Quantitative Variable:
'IMDb.Rating'

```{r}
summary(my_file$IMDb.Rating)
## head(dFile)
```

## 3. (3p) Data Visualization:

Use at least 4 useful, meaningful and different "data visualization
techniques" which will help you understand your data further
(distribution, outliers, variability, etc). Use 2 of the visualizations
to compare two groups (like female/male; smoker/non-smoker etc).

```{r} 

# Rename the columns to match the dataset
colnames(my_data) <- c("Const", "Your_Rating", "Date_Rated", "Title", "URL", "Title_Type", "IMDb_Rating", "Runtime_mins", "Year", "Genres", "Num_Votes", "Release_Date", "Directors")

# Data visualization techniques

# 1. Histogram of IMDb Ratings
ggplot(my_data, aes(x = IMDb_Rating)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Distribution of IMDb Ratings", x = "IMDb Rating", y = "Count") +
  theme_minimal()

# 2. Boxplot of Runtime by Title Type
ggplot(my_data, aes(x = Title_Type, y = Runtime_mins, fill = Title_Type)) +
  geom_boxplot() +
  labs(title = "Variability in Runtime by Title Type", x = "Title Type", y = "Runtime (mins)") +
  theme_minimal()

# 3. Barplot of Number of Votes by Year
ggplot(my_data, aes(x = factor(Year), y = Num_Votes, fill = factor(Year))) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Votes by Year", x = "Year", y = "Number of Votes") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(expand = c(0.01, 0.01))



# 4. Scatterplot of IMDb Ratings vs. Num Votes
ggplot(my_data, aes(x = IMDb_Rating, y = Num_Votes)) +
  geom_point(color = "blue") +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 1)) +
  scale_y_continuous(labels = comma) +
  labs(title = "IMDb Ratings vs. Number of Votes", x = "IMDb Rating", y = "Number of Votes") +
  theme_minimal()



```

## 4. (3p) Confidence Intervals:

Build '2 Confidence Intervals' step by step: Calculate the mean, then
standard error, and then the CI. Make "clear comments" about your
findings

```         
```

## 5. (3p) Transformation:

Implement one transformation (log transformation, Box-Cok
transformation, etc) for one of your quantitative variables, which is
not normally distributed; but will be normal or more normal, after the
transformation

```         
```

## 6. (2p every item if not indicated) t-test (Welch t-test or Wilcoxon rank-sum test or Paired t-test)

 Implement a single t-test for one of your "normally or
not-normally distributed" variable: a. Aim In words, what is your
objective here? b. Hypothesis and level of significance: Write your
hypothesis in scientific form and determine the level of singnificance.
c. Assumption Check (4p): Is your data independent or dependent? Tell
why you chose this test. Check the required assumptions statistically
and "comment on each of them is a must!". d. Indicate "which test you
choose" "for what reason" e. Result: Give the output of the test and
write down the result (ex: since p value is less /greater than alpha, I
reject/not reject the null hypothesis). f. Conclusion (4p): You got your
result in item e. Write down the conclusion of your result, in such a
way that, the reader who doesn't know any statistics can understand your
findings. g. What can be Type-1 and Type-2 error here? Not definition!
Tell these situations in terms of your data. (4p

```         
```

## 7. (2p every item if not indicated) Fisher's exact test for count data

a.  Aim In words, what is your objective? Provide the contingency table
    here.
b.  Hypothesis and level of significance: Write your hypothesis in
    scientific form and determine the level of singnificance.
c.  Result: Give the output of the test and write down the result (ex:
    since p value is less /greater than alpha, I reject/not reject the
    null hypothesis).
d.  Conclusion (4p): You got your result in item c. Write down the
    conclusion of your result, in such a way that, the reader who
    doesn't know any statistics can understand your findings.
e.  Odds Ratio (4p): Comment about the odds ratio, what does it
    indicate?

```         
```

## 8. (2p every item if not indicated) ANOVA and Tukey Test

a.  Aim In words, what is your objective here?
b.  Hypothesis and level of significance: Choose more than 2 (≥3) groups
    to compare! Write your hypothesis in scientific form and determine
    the level of singnificance.
c.  Assumption Check: Check the required assumptions statistically.
    "comment on each of them is a must!".
d.  Result of ANOVA: Give the output of the test and write down the
    result (ex:since p value is less than alpha, I reject the null
    hypothesis)
e.  Conclusion of ANOVA (4p): You got your result in item d. Write down
    the conclusion of your result, in such a way that, the reader who
    doesn't know any statistics can understand your findings.
f.  Result of Tukey: Give the output of the test and write down the
    result (ex: since p value is less /greater than alpha, I reject/not
    reject the null hypothesis)
g.  Conclusion of Tukey (4p): You got your result in item f. Write down
    the conclusion of your result, in such a way that, the reader who
    doesn't know any statistics can understand your findings.

```         
```

## 9. (2p every item) Multiple Linear Regression

a.  Aim In words, what is your objective here? Not definition, talk
    about your own aim/problem.
b.  Regression Equation: Multiple linear regression (MLR) is a
    statistical technique that uses several explanatory variables to
    predict the outcome of a response variable. Which ones are your
    explanatory variables and which one is your response variable? Write
    down the "statistical/mathematical equation" of your regression
    function using those variables and the parameters.
c.  Hypothesis and level of significance: Write your hypothesis in
    scientific form and determine the level of singnificance.
d.  Find the Best Model: Use step function and find the best model,
    describe the reason which makes it the best one.
e.  Assumption Check (4p): Check the required assumptions statistically,
    "comment on each of them is a must!".
f.  Result: Give the output of the best model and write down the result.
g.  Conclusion (4p): You got your result in item f. Write down the
    conclusion of your result, in such a way that, the reader who
    doesn't know any statistics can understand your findings.

```         
```
